Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-15 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
working on parspec section; new refs

We start by reviewing the parametrized ringdown spin expansion coefficients formalism (ParSpec)
introduced in Ref.~\cite{Maselli:2019mjd}, but following some of the notation of Ref.~\cite{Carullo:2021dui}.
%
The formalism modifies the Kerr quasinormal mode frequency $\omega_{J}^{\rm K}$ and
damping time $\tau_{J}^{\rm K}$ as
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{J} &= \omega^{\rm K}_{J} \, (1 + \delta \omega_{J}), \\
\tau_{J}   &= \tau^{\rm K}_{J}   \, (1 + \delta \tau_{J}),
\end{align}
\end{subequations}
%
where the $J$ collectively represents the QNM labels $\{l,\, m,\, n\}$, where $l$ is
the , $m$ and $n$ the ove

We expand the Kerr expression as:
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{J} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{J} \, \left( 1 + \gamma \delta \omega^{(n)}_{J} \right), \\
\tau_{J}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{J}   \, \left( 1 + \gamma \delta \tau^{(n)}_{J} \right),
\end{align}
\end{subequations}
%
where $M$ and $\chi$ are the detector-frame mass and dimensionless spin of the
remnant black hole; $\omega_{K}^{n}$, $\tau_{K}^{n}$ are dimensionless coefficients
of the spin expansion for Kerr QNMs; $\delta\omega_{K}^{n}$, $\delta\tau_{K}^{n}$ are
source-independent dimensionless coefficients that characterize the corrections to the Kerr QNM at each spin-order;
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p}
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of lenght, which sets the
lenghtscale in which beyond-GR modification become important and the exponent $p$
which is related to how the beyond-GR modifications are included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also in terms of the detector-frame mass $M$ through
a the redshift $z$.

In principle, a modification to GR would also affect $M$ and $\chi$ and the expansion should be written
in terms of non-GR mass and spin, say $\bar{M}$ and $\bar{\chi}$. If we assume that the beyond-GR corrections
are included perturbatively, the modifications to the BH mass and spin can be absorved into the deviations
parameters $\delta\omega$ and $\delta\tau$.
%
This means we can identify the $M$ and $\chi$ with their corresponding GR values.

>>>>

We start by reviewing the parametrized ringdown spin expansion coefficients
formalism (ParSpec) introduced in Ref.~\cite{Maselli:2019mjd}, but following
some of the notation of Ref.~\cite{Carullo:2021dui}.
%
A general procedure to describe deviations from the Kerr quasinormal mode frequencies $\omega_{J}^{\rm K}$ and
damping time $\tau_{J}^{\rm K}$ is to write them as,
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{J} &= \omega^{\rm K}_{J} \, (1 + \delta \omega_{J}), \\
\tau_{J}   &= \tau^{\rm K}_{J}   \, (1 + \delta \tau_{J}),
\end{align}
\end{subequations}
%
as done e.g., in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Carullo:2018sfu}.
%
Here we use $J$ to collectively represents the QNM labels $\{l,\, m,\, n\}$,
where $l$ and $m$ are the multipole indices, and $n$ the overtone number.
%
We work exclusively with $\ell = m = 2$ and $n = 0$ (fundamental mode) in this work.

We expand the Kerr expression as:
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{J} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{J} \, \left( 1 + \gamma \delta \omega^{(n)}_{J} \right), \\
\tau_{J}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{J}   \, \left( 1 + \gamma \delta \tau^{(n)}_{J} \right),
\end{align}
\end{subequations}
%
where $M$ and $\chi$ are the detector-frame mass and dimensionless spin of the
black hole; $\omega_{K}^{n}$, $\tau_{K}^{n}$ are dimensionless coefficients
of the spin expansion for Kerr QNMs; $\delta\omega_{K}^{n}$, $\delta\tau_{K}^{n}$ are
source-independent dimensionless coefficients that characterize the corrections to the Kerr QNM at each spin-order;
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$ given by,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of length (and sets the
lengthscale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by using the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}.

In principle, a modification to GR would also affect $M$ and $\chi$ and the expansion should be written
in terms of non-GR mass and spin, say $\bar{M}$ and $\bar{\chi}$. If we assume that the beyond-GR corrections
are included perturbatively, the modifications to the BH mass and spin can be absorved into the deviations
parameters $\delta\omega$ and $\delta\tau$.
%
This means we can identify the $M$ and $\chi$ with their corresponding GR values.
%
